---
title: "Package"
author: "Lucie Contamin"
date: "6/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
Introduction to developing R package and R project:

* To be organized
* Reusable

Two different ways of organising codes: project (less constraints) package. 

* Project is a specific folder, the work is saved, maintained the status of work (linking codes to data), always return to previous.

* Package: with architecture (with rules), allow to publish to public repo. Can be code and/or data.
Repo: cran, github, bioconductor etc.

## Demo


```{r, eval=T}
# Make an example data frame
lab_res <- data.frame("test1" = c(1, 3),
           "test2" = c(3, 4),
           "ID" = c("a", "b"))
print(lab_res)

# write the data frame in a file in the working directory
write.csv(lab_res, file = paste0(getwd(),"/lab_res.csv"))

# come back and read
lab <- read.csv("lab_res.csv")

# to print a variable you can do
lab
# or
print(lab)

# or print along with assignment
(lab <- read.csv("lab_res.csv"))

# let assume we need to make a function

mean_col <- function(x){
    tot <- sum(x)/length(x)
    return(tot)
}

(m_test1 <- mean_col(lab$test1))

```

Let say we are happy with out script, we leave and come back 3 years laters. We don't know what was going on. Without documentation, we forget what the scripts are meant to. So, we organise in project.
Use the menu of RStudio. What happened was:

* A project file is created: .Rproj, create a folder R for scripts, data for data. 
Another way of making project is to use usethis::create_project (with R folder created automatically to store scripts)

```{r, eval=FALSE}
usethis::create_project("/Users/haihb/Desktop/cafer")
```
Easy to share. But then we'll need to add documentation to our script to reuse/share: Add comments with # (like above).

Now you want to make a package:
* From interface
* using usethis
```{r, eval=F}
usethis::create_package("/Users/haihb/Desktop/cafer1")
```

Pay attention to the architecture created. You can put your code in R. For now, let's play with hello.R

Start describing your package with DESCRIPTION. 

Make data folder with usethis

```{r, eval=F}
# This function create data-raw folder and an empty dataset
usethis::use_data_raw()
usethis::use_data("DATASET")
# put in the data-raww the script to create data, in DATASET.R

lab_res <- data.frame("test1" = c(1, 3),
                      "test2" = c(3, 4),
                      "ID" = c("a", "b"))

usethis::use_data(lab_res)

```

To document the data objects. Create in R, the file data.R. You need to document the data, if not warnings while compile. (Roxygen, language for documentation).

Use `devtools::document()` to create the help files in the man folder.

Now we go back to document the functions in scripts (functions). Create a script file for funtion add

```{r}
add <- function(x, y = 1){
    if(is.character(x)){
        stop("x should be a number")
    }
    x <- x + y
    return(x)
}
```

Then save it with the same name of your function: add.R in R folder. 

To document a function, it requires:

* @params to describe the parameters. Compulsory.
* @export if the function is loaded in the env to share. We don't need to export all functions but the shared ones needed to be exported.

```{r, eval=F}

#' Addition
#' 
#' Calculate the sum
#' @param x a numeric value
#' @export
#' @examples
#' x <- 1
#' add(x)
#'

add <- function(x, y = 1){
    if(is.character(x)){
        stop("x should be a number")
    }
    x <- x + y
    return(x)
}

```

